---
title: "NFL"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation Analysis to decide the winner between two NFL teams


```{r echo=FALSE,warning=FALSE}
install.packages("XML",repos="http://cran.rstudio.com/")
install.packages("RSQLite",repos="http://cran.rstudio.com/")
install.packages("stringr",repos="http://cran.rstudio.com/")
install.packages("ggplot2",repos="http://cran.rstudio.com/")
library(XML)
library(RSQLite)
library(stringr)
library(ggplot2)
year <- 2015
url <- paste("http://sports.yahoo.com/nfl/stats/byteam?group=Offense&cat=Total&conference=NFL&year=season_",year,"&sort=530&old_category=Total&old_group=Offense")
offense <- readHTMLTable(url,colClasses ="character",encoding="UTF-8")[[7]]
head(offense)
```

Step 2: After inspecting the table, we see there are some blank columns plus the classes of these columns are factors. We need to change that
```{r}
offense <- offense[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26,28)]
offense[,1] <- as.character(offense[,1])
offense[,2:13] <- apply(offense[,2:13],2,as.numeric)
offense[,14] <- as.numeric(substr(offense[,14], 1, 2))*60 + as.numeric(substr(offense[,14], 4, 6))
head(offense)
```

Next step in this project would be to build up a roaster and run the simulation to decide the winner.

Step 3: Now we will repeat the steps for defense
```{r}
url <- paste("http://sports.yahoo.com/nfl/stats/byteam?group=Defense&cat=Total&conference=NFL&year=season_",year,"&sort=530&old_category=Total&old_group=Defense")
defense <- readHTMLTable(url, encoding = "UTF-8",colClasses="character")[[7]]
defense <- defense[,-c(2,4,6,8,10,12,14,16,18,20,22,24,26)]
defense[,1] <- as.character(defense[,1])
defense[,2:13] <- apply(defense[,2:13],2,as.numeric)
combined <- merge(offense, defense, by.x="Team", by.y="Team")
colnames(combined)[2] <- "Games"
colnames(combined)[3] <- "OffPPG"
colnames(combined)[4] <- "OffYPG"
colnames(combined)[5] <- "OffPassYPG"
colnames(combined)[6] <- "OffRushYPG"
combined$G.y <- NULL
colnames(combined)[15] <- "DefPPG"
colnames(combined)[16] <- "DefYPG"
colnames(combined)[17] <- "DefRushYPG"
colnames(combined)[18] <- "DefPassYPG"
```
## Including Plots



```{r }
Off_PPG <- hist(combined$OffPPG, breaks=10, main="Offensive Points Per Game", xlab="Offensive PPG",ylab="Number of Teams")
Def_YPG <- hist(combined$DefYPG, breaks=10, main="Defensive Points Per Game", xlab="Defense PPG",ylab="Number of Teams")
```

Step 4 : Convert table to Dataframe by using the transform function
```{r }
ppg <- transform(combined,Team=reorder(Team,combined$OffPPG))
ypg <- transform(combined,Team=reorder(Team,-combined$DefYPG))
```

Step 5: Compare Points per game by Team for both Offense and Defense
```{r }
ggplot(ppg,aes(x=Team,y=OffPPG))+geom_bar(stat="Identity",color="darkgreen",fill="white")+coord_flip()+labs(x="Team",y="Avg Points per game")+ggtitle("Avg points per game")

ggplot(ypg,aes(x=Team,y=DefYPG))+geom_bar(stat="Identity",color="darkgreen",fill="white")+coord_flip()+labs(x="Team",y="Avg Points per game")+ggtitle("Avg points per game")
```

From these charts, we can get a visual sense osf what fans saw throughout the season, specifically the incredible offense of the Denver Broncos team and the unstoppable defense of the Seattle Seahawks team, which ultimately led them to a Super Bowl victory against the Broncos.

```{r }
ggplot(combined, aes(x=combined$OffYPG, y=combined$OffPPG)) +
  geom_point(shape=5, size=2) + geom_smooth() + 
  labs(x="Yards per Game",y="Points per Game") + ggtitle("Offense Yards vs. Points per Game") + 
  theme(plot.title = element_text(size=18, face="bold"))

ggplot(combined, aes(x=combined$DefYPG, y=combined$DefPPG)) +
  geom_point(shape=5, size=2) + geom_smooth() + 
  labs(x="Yards Allowed per Game",y="Points Alloed per Game") + ggtitle("Defense Yards vs. Points per Game") + 
  theme(plot.title = element_text(size=18, face="bold"))
```

As you can see, these two variables are positively correlated—as yards per game increases, points per game also usually increases.We calculate correlation now

```{r }
cor(combined$OffYPG,combined$OffPPG)
cor(combined$DefYPG,combined$DefPPG)
```

We will now use team level statistics to construct indexes for the offensive and defensive strengths of each team. The offense strength index will depend on the teams' passing and rushing strength, and the defense strength index will depend on the ability of the teams to defend against the pass and the rush. This will allow us compare the different aspects of each team's game to other teams and will let us arrive at a winner and a loser in our simulated games later on.

```{r }
offense$OPassStrength <- max(offense[,5])-offense[,5]
offense$OPassStrength <- (1-(offense$OPassStrength/max(offense$OPassStrength)))*100
offense$ORushStrength <- max(offense[,6])-offense[,6]
offense$ORushStrength <- (1-(offense$ORushStrength/max(offense$ORushStrength)))*100
offense$OPPGStrength <- max(offense[,3])-offense[,3]
offense$OPPGStrength <- (1-(offense$OPPGStrength/max(offense$OPPGStrength)))*100

offense$OYPGStrength <- max(offense[,4])-offense[,4]
offense$OYPGStrength <- (1-(offense$OYPGStrength/max(offense$OYPGStrength)))*100
offense$OffStrength <- (offense$OPassStrength+offense$ORushStrength+offense$OPPGStrength+offense$OYPGStrength)/4

defense$DPassStrength <- max(defense[,6])-defense[,6]
defense$DPassStrength <- defense$DPassStrength/max(defense$DPassStrength)*100
defense$DRushStrength <- max(defense[,5])-defense[,5]
defense$DRushStrength <- defense$DRushStrength/max(defense$DRushStrength)*100

defense$DPPGStrength <- max(defense[,3])-defense[,3]
defense$DPPGStrength <- defense$DPPGStrength/max(defense$DPPGStrength)*100

defense$DYPGStrength <- max(defense[,4])-defense[,4]
defense$DYPGStrength <- defense$DYPGStrength/max(defense$DYPGStrength)*100

defense$DefStrength <- (defense$DPassStrength+defense$DRushStrength+defense$DPPGStrength+defense$DYPGStrength)/4
```

Step 6: Starting with Game Simulator
```{r }
home_team <- "Chicago Bears"
away_team <- "New Orleans Saints"

off_game <- subset(offense,Team==home_team | Team==away_team)[,c(1,15,16,19)]
def_game <- subset(defense,Team==home_team | Team==away_team)[,c(1,14,15,18)]
game <- merge(off_game,def_game,by.x="Team",by.y="Team")
game$Net_Pass[game$Team==home_team] <- game$OPassStrength[game$Team==home_team] - game$DPassStrength[game$Team==away_team]
game$Net_Pass[game$Team==away_team] <- game$OPassStrength[game$Team==away_team] - game$DPassStrength[game$Team==home_team]

game$Net_Rush[game$Team==home_team] <- game$ORushStrength[game$Team==home_team] - game$DRushStrength[game$Team==away_team]
game$Net_Rush[game$Team==away_team] <- game$ORushStrength[game$Team==away_team] - game$DRushStrength[game$Team==home_team]
game$Net_Total[game$Team==home_team] <- game$OffStrength[game$Team==home_team] - game$DefStrength[game$Team==away_team]
game$Net_Total[game$Team==away_team] <- game$OffStrength[game$Team==away_team] - game$DefStrength[game$Team==home_team]

game$Net_Total <- game$Net_Pass + game$Net_Rush + game$Net_Total

if(game$Net_Total[game$Team==home_team] >= game$Net_Total[game$Team==away_team]){
  winner <- home_team
  loser <- away_team
}else{
  winner <- away_team
  loser <- home_team
}

print(paste(winner, "beat", loser)) 
```